---
title: "Lecture 2 - Data Storage"
output: html_notebook
---

```{r}
library(readr)
raw <- read_csv("lecture_02_data.csv")
```

* Zero encoded as a dash and is in Flag column
* Flag column represents Unavailable
* Data (age range) stored in variable names
* Unusual characters in city and variable names
* Mix of aggregate data with raw data (region totals and overall total)
* Corresponding region not included with city
    - New England: MA, CT, RI
    - Mid. Atlantic: NY, PA, NJ
    - E.N. Central: OH, IL, MI, IN, WI
    - W.N. Central: IA, MN, KS, MO, NE; Lincoln, NC?
    - S. Atlantic: GA, MD, NC, FL, VA, D.C., DE
    - E.S. Central: AL, TN, KY
    - W.S. Central: TX, LA, AR, OK
    - Mountain: NM, ID, CO, NV, UT, AZ
    - Pacific: CA, HI, OR, WA

```{r}
library(dplyr)
library(tidyr)
library(stringr)

tidy <- raw %>%
    gather(age_range, mortalities, `All causes, by age (years), All Ages`:`P&I† Total, flag`) %>%
    mutate(`Reporting Area` = iconv(`Reporting Area`, "latin1", "ASCII", sub = ""),
           age_range = iconv(age_range, "latin1", "ASCII", sub = "-"),
           age_range = str_replace_all(age_range, "---", "-"),
           age_range = str_replace_all(age_range, "All causes, by age \\(years\\),", ""),
           age_range = str_replace_all(age_range, ", flag", ""),
           age_range = str_replace_all(age_range, "P&I- Total", "pna_inf"),
           mortalities = str_replace_all(mortalities, "-", 0),
           `Reporting Area` = str_replace_all(`Reporting Area`, "Lincoln, NC", "Lincoln, NE")) %>%
    drop_na(mortalities) %>%
    arrange(`Reporting Area`, `MMWR YEAR`, `MMWR WEEK`, age_range) %>%
    select(-`Location 2`) %>%
    filter(!`Reporting Area` %in% c("New England", "Mid. Atlantic", "E.N. Central", "W.N Central", "S. Atlantic", "E.S. Central", "W.S. Central", "Mountain", "Pacific", "Total"))

tidy$region <- case_when(str_detect(tidy$`Reporting Area`, "MA|CT|RI") ~ "New England",
                         str_detect(tidy$`Reporting Area`, "NY|PA|NJ") ~ "Mid. Atlantic",
                         str_detect(tidy$`Reporting Area`, "OH|IL|MI|IN|WI") ~ "E.N. Central",
                         str_detect(tidy$`Reporting Area`, "IA|MN|KS|MO|NE") ~ "W.N. Central",
                         str_detect(tidy$`Reporting Area`, "GA|MD|NC|FL|VA|D.C.|DE") ~ "S. Atlantic",
                         str_detect(tidy$`Reporting Area`, "AL|TN|KY") ~ "E.S. Central",
                         str_detect(tidy$`Reporting Area`, "TX|LA|AR|OK") ~ "W.S. Central",
                         str_detect(tidy$`Reporting Area`, "NM|ID|CO|NV|UT|AZ") ~ "Mountain",
                         str_detect(tidy$`Reporting Area`, "CA|HI|OR|WA") ~ "Pacific")    
```

1. In Houston, which week had the largest number of mortalities among patients 45 years and older?
1. What city in the US had the largest number of mortalities in a given week among patients 45 years and older?
1. What is the average number of mortalities among patients 45 years and older in cities within the W.S. Central region?
