---
title: "Lecture 2 - Data Storage"
output: html_notebook
---

```{r}
library(readr)
raw <- read_csv("lecture_02_data.csv")
```

* Zero encoded as a dash and is in Flag column
* Flag column represents Unavailable
* Data (age range) stored in variable names
* Unusual characters in city and variable names
* Mix of aggregate data with raw data (region totals and overall total)
* Corresponding region not included with city
    - New England: MA, CT, RI
    - Mid. Atlantic: NY, PA, NJ
    - E.N. Central: OH, IL, MI, IN, WI
    - W.N. Central: IA, MN, KS, MO, NE; Lincoln, NC?
    - S. Atlantic: GA, MD, NC, FL, VA, D.C., DE
    - E.S. Central: AL, TN, KY
    - W.S. Central: TX, LA, AR, OK
    - Mountain: NM, ID, CO, NV, UT, AZ
    - Pacific: CA, HI, OR, WA

```{r}
library(dplyr)
library(tidyr)
library(stringr)

tidy <- raw %>%
    gather(age_range, mortalities, `All causes, by age (years), All Ages`:`P&I† Total, flag`) %>%
    mutate(age_range = iconv(age_range, "latin1", "ASCII", sub = "-"),
           age_range = str_replace_all(age_range, "---", "-"),
           age_range = str_replace_all(age_range, "All causes, by age \\(years\\),", ""))
    
```

1. In Houston, which week had the largest number of mortalities among patients 45 years and older?
1. What city in the US had the largest number of mortalities in a given week among patients 45 years and older?
